---
title: "Statistical Inference Course Project Part 2"
author: "Gaurab Kundu"
date: "2022-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 2: Basic Inferential Data Analysis Instructions

## Overview

In this second part, we perform basic inferential analyses using the ToothGrowth data in the R datasets package.

As per the help file of this dataset, the response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

## Load Necessary Libraries

```{r}
library(ggplot2)
```

## Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r}
# load the dataset
library(datasets)
data(ToothGrowth)
# look at the dataset variables
str(ToothGrowth)
```

```{r}
# convert variable dose from numeric to factor
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
# look at the dataset variables after conversion
str(ToothGrowth)
```

## Provide a basic summary of the data.

```{r}
# summary statistics for all variables
summary(ToothGrowth)
```

```{r}
# split of cases between different dose levels and delivery methods
table(ToothGrowth$dose, ToothGrowth$supp)
```

