---
title: "Statistical Inference Course Project Part 1"
author: "Gaurab Kundu"
date: "2022-10-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1: Simulation Exercise

## Overview

In this part of the project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. I will investigate the distribution of averages of 40 exponentials. I will be doing a thousand simulations.

I will iIllustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials.

## Questions

1. Show the sample mean and compare it to the theoretical mean of the distribution.

2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

3. Show that the distribution is approximately normal.

In point 3, I will focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

## Simulations

It is good practice to set the seed of R’s random number generator before simulating random observations so that the simulations can be reproduced.

```{r}
set.seed(1234)
```

Set the rate parameter (lambda) equal to 0.2, set the number of exponentials equal to 40, and set the number of simulations equal to 1000 as instructed in the rubric.

```{r}
lambda <- 0.2
n <- 40
sim <- 1000
```

Simulate the 40 exponentials.

```{r}
simExp <- replicate(sim, rexp(n, lambda))
```

Calculate the mean of the simulated exponentials.

```{r}
meanExp <- apply(simExp, 2, mean)
```

